---
title: "Shiny"
author: "Chang Lu"
date: "2024-11-16"
output: html_document
runtime: shiny
---


## Your First Shiny App

This document demonstrates two Shiny apps:
1. A **Text Input App** that takes a user's name and displays a greeting.
2. A **Slider Input App** that multiplies a slider value by 5 and displays the result.
3. A **Caluculator App** that displays the computed result in real-time.
4. A **Dynamic Multiplication and Adjustment Calculator App** that dynamically calculates the product of two inputs amd adjusts the product by adding fixed values.
5. A **Interactive Data Exploration App** that users can select a dataset from a predefined list (economics, faithfuld, or seals), all of which are included in the ggplot2 package. Then view a summary of the selected dataset, providing key statistical insights and visualize the dataset as a plot, dynamically generated based on the their selection.

Below, you can see the code and interactive examples for each app, as well as their differences.

---

### Example 1: Text Input App


```{r}
library(shiny)

ui <- fluidPage(
    textInput("name", "What's your name?"),
    textOutput("greeting")
)

server <- function(input, output, session) {
    output$greeting <- renderText({
        paste0("Hello ", input$name)
    })
}

shinyApp(ui, server)
```

### Example 2: Slider Input App

```{r}
ui <- fluidPage(
    sliderInput("x", label = "If x is", min = 1, max = 50, value = 30),
    "then x times 5 is",
    textOutput("product")
)

server <- function(input, output, session) {
    output$product <- renderText({ 
        input$x * 5
    })
}

shinyApp(ui, server)
```

### Example 3: Calculator App
```{r}
ui <- fluidPage(
    sliderInput("x", label = "If x is", min = 1, max = 50, value = 30),
    sliderInput("y", label = "and y is", min = 1, max = 50, value = 30),
    "then x times y is",
    textOutput("product")
)

server <- function(input, output, session) {
    output$product <- renderText({ 
       input$x * input$y
    })
}

shinyApp(ui, server)
```

### Example 4: Dynamic Multiplication and Adjustment Calculator App
```{r}

ui <- fluidPage(
    sliderInput("x", "If x is", min = 1, max = 50, value = 30),
    sliderInput("y", "and y is", min = 1, max = 50, value = 5),
    "then, (x * y) is", textOutput("product"),
    "and, (x * y) + 5 is", textOutput("product_plus5"),
    "and (x * y) + 10 is", textOutput("product_plus10")
)

server <- function(input, output, session) {
    # Add this reactive expression to reduce 
    # amount of duplicated code
    product <- reactive({
        input$x * input$y
    })
    output$product <- renderText({ 
        product()
    })
    output$product_plus5 <- renderText({ 
        product() + 5
    })
    output$product_plus10 <- renderText({ 
        product() + 10
    })
}


shinyApp(ui, server)
```

### Example 5: Interactive Data Exploration App
```{r}
library(ggplot2)

datasets <- c("economics", "faithfuld", "seals")

ui <- fluidPage(
    selectInput("dataset", "Dataset", choices = datasets),
    verbatimTextOutput("summary"),
    # 1st Bug: tableOutput -> plotOutput
    plotOutput("plot")
)

server <- function(input, output, session) {
    dataset <- reactive({
        get(input$dataset, "package:ggplot2")
    })
    # 2nd Bug: Spelling
    output$summary <- renderPrint({
        summary(dataset())
    })
    output$plot <- renderPlot({
      # 3rd Bug: dataset -> dataset() 
        plot(dataset())
    }, res = 96)
}

shinyApp(ui, server)
```

## Exercise 2.3.5

### 1

a. renderPrint(summary(mtcars)) with verbatimTextOutput
Correct: Yes
Reason: summary(mtcars) produces a detailed, console-like output (e.g., summary statistics), which requires verbatimTextOutput to preserve the formatting.

b. renderText("Good morning!") with textOutput
Correct: Yes
Reason: "Good morning!" is simple text without special formatting, so textOutput is appropriate.

c. renderPrint(t.test(1:5, 2:6)) with verbatimTextOutput
Correct: Yes
Reason: The t.test output is a structured result of a statistical test, displayed in console-like formatting, which is correctly handled by verbatimTextOutput.

d. renderText(str(lm(mpg ~ wt, data = mtcars))) with verbatimTextOutput
Correct: No
Reason: While renderText is used, the str() function's output is console-like and is better suited for renderPrint paired with verbatimTextOutput. Using renderText here would lead to improper formatting of the output.

### 2

```{r}
ui <- fluidPage(
    sliderInput("n", "Number of Points", min = 5, max = 20, value = 5),
    plotOutput("plot", width = "700px", height = "300px"),
    p("This is a scatterplot of numbers from 1 to n(as you choose).")
)

server <- function(input, output, session) {
    output$plot <- renderPlot({
        plot(1:input$n)
    }, res = 96)
}

shinyApp(ui, server)
```

### 3
```{r}
library(DT)

ui <- fluidPage(
    titlePanel("Interactive mtcars Data Table"),
    
    # Row selection input
    sidebarLayout(
        sidebarPanel(
            sliderInput("rows", "Rows per page:", min = 5, max = 20, value = 5, step = 5),
            selectInput("cyl_filter", "Filter by number of cylinders:",
                        choices = c("All", unique(mtcars$cyl)), selected = "All"),
            downloadButton("downloadData", "Download Table")
        ),
        
        mainPanel(
            dataTableOutput("table")
        )
    )
)

server <- function(input, output, session) {
    # Reactive expression to filter data based on selected cylinders
    filteredData <- reactive({
        if (input$cyl_filter == "All") {
            mtcars
        } else {
            subset(mtcars, cyl == as.numeric(input$cyl_filter))
        }
    })
    
    # Render the DataTable with reactive options
    output$table <- renderDataTable({
        datatable(filteredData(),
                  options = list(pageLength = input$rows,
                                 ordering = TRUE,
                                 searching = TRUE))
    })
    
    # Download handler for exporting the data
    output$downloadData <- downloadHandler(
        filename = function() { "filtered_mtcars.csv" },
        content = function(file) {
            write.csv(filteredData(), file)
        }
    )
}

shinyApp(ui, server)
```

### 4
```{r}
library(reactable)
library(dplyr)

ui <- fluidPage(
  titlePanel("Interactive mtcars Table"),
  sidebarLayout(
    sidebarPanel(
      selectInput("cyl_filter", "Filter by Cylinders:",
                  choices = c("All", unique(mtcars$cyl)), selected = "All"),
      downloadButton("downloadData", "Download Table")
    ),
    mainPanel(
      reactableOutput("table")
    )
  )
)

server <- function(input, output, session) {
  # Reactive expression to filter data based on user input
  filteredData <- reactive({
    if (input$cyl_filter == "All") {
      mtcars
    } else {
      mtcars %>% filter(cyl == as.numeric(input$cyl_filter))
    }
  })
  
  # Render the interactive table
  output$table <- renderReactable({
    reactable(
      filteredData(),
      searchable = TRUE,    # Enable search
      striped = TRUE,       # Add striped rows for readability
      highlight = TRUE,     # Highlight rows on hover
      bordered = TRUE,      # Add borders to the table
      defaultPageSize = 10, # Show 10 rows per page
      columns = list(
        mpg = colDef(name = "Miles/Gallon", 
                     style = function(value) {
                       if (value > 20) list(color = "green", fontWeight = "bold") else NULL
                     }),
        cyl = colDef(name = "Cylinders", align = "center"),
        hp = colDef(name = "Horsepower")
      )
    )
  })
  
  # Add a download handler for exporting the filtered data
  output$downloadData <- downloadHandler(
    filename = function() { "filtered_mtcars.csv" },
    content = function(file) {
      write.csv(filteredData(), file)
    }
  )
}

shinyApp(ui, server)
```

## Exercise 3.3.6

### 1
```{r}
server1 <- function(input, output, session) {
    output$greeting <- renderText({
        paste0("Hello ", input$name)
    })
}
```

Explanation: Fixes the issue by correctly using renderText and referencing input$name.

### 2 
```{r}
server2 <- function(input, output, session) {
    greeting <- reactive({
        paste0("Hello ", input$name)
    })
    output$greeting <- renderText({
        greeting()
    })
}
```

Explanation: 
1. Introduces a reactive expression greeting to handle the text construction separately.
2. Requires parentheses when using the reactive greeting().

### 3
```{r}
server3 <- function(input, output, session) {
    output$greeting <- renderText({
        paste0("Hello ", input$name)
    })
}
```
Explanation: Fixes a spelling error (output$greeting instead of output$greting) and ensures renderText is used properly.

## Exercise 4.8

### 2
```{r}
# downloading the data
dir.create("neiss")

download <- function(name) {
  url <- "https://github.com/hadley/mastering-shiny/raw/main/neiss/"
  download.file(paste0(url, name), paste0("neiss/", name), quiet = TRUE)
}

download("injuries.tsv.gz")
download("population.tsv")
download("products.tsv")
```
```{r}
library(forcats)
injuries <- vroom::vroom("neiss/injuries.tsv.gz")
injuries
```

```{r}
# flipped code
injuries %>%
  mutate(diag = fct_infreq(fct_lump(diag, n = 5))) %>%
  group_by(diag) %>%
  summarise(n = as.integer(sum(weight)))
```

### 3

#### Purpose

This Shiny app analyzes injury data (injuries.tsv.gz) based on user-selected product codes (products.tsv). It allows users to:

View summary tables of injuries categorized by diagnosis, body part, and location.
Visualize the relationship between age and injury count or rate by gender.
Explore specific injury narratives with a single click.

#### code
```{r}
library(dplyr)
library(ggplot2)
library(forcats)
library(vroom)
library(shiny)

# Load datasets
injuries <- vroom::vroom("neiss/injuries.tsv.gz")
products <- vroom::vroom("neiss/products.tsv")
population <- vroom::vroom("neiss/population.tsv")

# Custom function to count and summarize top categories
count_top <- function(df, var, n = 5) {
  df %>%
    mutate({{ var }} := fct_lump(fct_infreq({{ var }}), n = n)) %>%
    group_by({{ var }}) %>%
    summarise(n = as.integer(sum(weight)), .groups = "drop")
}

# Define UI
ui <- fluidPage(
  titlePanel("Injury Data Explorer"),
  fluidRow(
    column(8,
           selectInput("code", "Product",
                       choices = setNames(products$prod_code, products$title),
                       width = "100%")
    ),
    column(2, selectInput("y", "Y axis", c("rate", "count"))),
    column(2, numericInput("num_rows", "Number of Rows", value = 5, min = 0, max = 10))
  ),
  fluidRow(
    column(4, tableOutput("diag")),
    column(4, tableOutput("body_part")),
    column(4, tableOutput("location"))
  ),
  fluidRow(
    column(12, plotOutput("age_sex"))
  ),
  fluidRow(
    column(2, actionButton("story", "Tell me a story")),
    column(8, textOutput("narrative")),
    column(2, downloadButton("download", "Download Data"))
  )
)

# Define server logic
server <- function(input, output, session) {
  # Dynamically update product dropdown
  observe({
    updateSelectInput(session, "code", choices = setNames(products$prod_code, products$title))
  })
  
  # Reactive dataset filtered by product code
  selected <- reactive({
    req(input$code)  # Ensure valid input
    injuries %>% filter(prod_code == input$code)
  })
  
  # Render summary tables with input validation
  output$diag <- renderTable({
    data <- count_top(selected(), diag)
    req(input$num_rows <= nrow(data))  # Validate number of rows
    data %>% slice(1:input$num_rows)
  }, width = "100%")
  
  output$body_part <- renderTable({
    data <- count_top(selected(), body_part)
    req(input$num_rows <= nrow(data))
    data %>% slice(1:input$num_rows)
  }, width = "100%")
  
  output$location <- renderTable({
    data <- count_top(selected(), location)
    req(input$num_rows <= nrow(data))
    data %>% slice(1:input$num_rows)
  }, width = "100%")
  
  # Generate summary for plotting
  summary <- reactive({
    selected() %>%
      count(age, sex, wt = weight) %>%
      left_join(population, by = c("age", "sex")) %>%
      mutate(rate = n / population * 1e4)
  })
  
  # Render plot
  output$age_sex <- renderPlot({
    req(summary())
    if (input$y == "count") {
      summary() %>%
        ggplot(aes(age, n, colour = sex)) +
        geom_line() +
        labs(y = "Estimated number of injuries")
    } else {
      summary() %>%
        ggplot(aes(age, rate, colour = sex)) +
        geom_line(na.rm = TRUE) +
        labs(y = "Injuries per 10,000 people")
    }
  }, res = 96)
  
  # Generate random narrative
  narrative_sample <- eventReactive(
    list(input$story, selected()),
    selected() %>% pull(narrative) %>% sample(1)
  )
  output$narrative <- renderText({
    req(narrative_sample())
    narrative_sample()
  })
  
  # Download handler for filtered data
  output$download <- downloadHandler(
    filename = function() {
      paste0("injury_data_", input$code, ".csv")
    },
    content = function(file) {
      write.csv(selected(), file, row.names = FALSE)
    }
  )
}

# Run the Shiny app
shinyApp(ui, server)
```